Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-06-07 PAPER.tex
syntax edits to 2.1.2

%First, 
We compute the power spectrum of our toy model dataset $\textbf{x}$ using Equations \ref{eq:qhat} and \ref{eq:phat}, with $\textbf{R} \equiv \widehat{\textbf{C}}^{-1}$.  Figure \ref{fig:toy_sigloss12} shows the estimated covariances of our toy model datasets along with $\widehat{\textbf{C}}^{-1}$ weighted data. The foreground sinusoid is clearly visible in $\widehat{\textbf{C}}_{\rm FG}$.  The power spectrum result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3}. Also plotted in the figure are the uniform-weighted ($\textbf{R} \equiv \textbf{I}$) power spectrum of the individual components $\textbf{x}_{\rm FG}$ (blue) and $\textbf{x}_{\rm EoR}$ (red). As shown, our $\widehat{\textbf{C}}^{-1}$ weighted result successfully suppresses foregrounds,
%For our toy model, the successful suppression of the foreground mode is 
demonstrated in Figure \ref{fig:toy_sigloss3} by the missing foreground peak in the weighted power spectrum estimate (green).  It is also evident that our result fails to recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low.  It is this behavior which we describe as signal loss.

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss3.pdf}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss4.pdf}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of foregrounds peaks at a $k$-mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare empirically estimated inverse covariance weighting where $\textbf{C}$ is derived from the data (left), and projecting out the first eigenmode only (right). In the former case, signal loss arises from the coupling of the eigenmodes of $\widehat{\textbf{C}}$ to the data. 
% JEA - redundant
%For an empirically estimated $\widehat{\textbf{C}}$, 
%its eigenvalues differ from those of the true covariance, where 
%the coupling between EoR-dominated eigenmodes and the data can lead to loss.
%Hence, 
There is negligible signal loss when the other eigenmodes besides the foreground one are no longer correlated with the data.
%assigning identical weights to all eigenmodes except the first, 
%constructing a covariance proportional to the identity and the outer product of the sinusoid eigenmode, since we are only down-weighting the FG-dominated mode in this case.
}
	\label{fig:toy_sigloss3}
\end{figure*}

As discussed in Section \ref{sec:QE}, this behavior is {\it not} expected in the case that we were to use $\textbf{C}^{-1}$ weighting.  Rather, we would obtain the behavior shown in the toy model in Appendix \ref{sec:icw_appendix}\footnote{Note that there the true covariance matrix is also the sum of a diagonal portion describing the signal, and a single mode describing the contaminant (c.f. Figure \ref{fig:toy_sigloss12}).}, with suppression of the foreground mode resulting in a nearly unbiased estimate of the power spectrum.   The key difference is that since $\widehat{\textbf{C}}$ is estimated from the data, its eigenvectors and eigenvalues are strongly coupled to the particular data realization that was used to compute it.

For the case of a mode which can be safely assumed to be predominantly a foreground, its presence in the covariance matrix will result in the desired suppression via a kind of projection (c.f. Equations \ref{eq:IdealToyModelCovariance} and \ref{eq:IdealToyModelEstimator});  whether or not it is strongly correlated with the the actual data vector is irrelevant.  However, in the case the covariance matrix is empirically estimated, the eigenmodes of $\textbf{C}_{\rm EoR}$ will both be incorrect and correlated with the data.  Incorrect eigenmodes of $\textbf{C}_{\rm EoR}$ that are not correlated with the data will lead to non-minimum variance estimates, but will not produce the suppression of the power spectrum amplitude seen.  As indicated in Appendix \ref{sec:sigloss_appendix}, however, having $\widehat{\mathbf{C}}$ correlated with the data vector $\mathbf{x}$ also produces a kind of projection of power in the {\it non}-foreground modes from the resulting power spectrum estimate, thus producing an estimate that is biased low.  In short, {\it if the covariance is computed from the data itself, it carries the risk of over-fitting information in the data and introducing a multiplicative bias (per $k$) to estimates of the signal.}

>>>>

%First, 
We compute the power spectrum of our toy model dataset $\textbf{x}$ using Equations \ref{eq:qhat} and \ref{eq:phat}, with $\textbf{R} \equiv \widehat{\textbf{C}}^{-1}$.  Figure \ref{fig:toy_sigloss12} shows the estimated covariances of our toy model datasets along with the $\widehat{\textbf{C}}^{-1}$ weighted data. The foreground sinusoid is clearly visible in $\widehat{\textbf{C}}_{\rm FG}$.  The power spectrum result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3}. Also plotted in the figure are the uniform-weighted ($\textbf{R} \equiv \textbf{I}$) power spectrum of the individual components $\textbf{x}_{\rm FG}$ (blue) and $\textbf{x}_{\rm EoR}$ (red). As shown, our $\widehat{\textbf{C}}^{-1}$ weighted result successfully suppresses foregrounds,
%For our toy model, the successful suppression of the foreground mode is 
demonstrated in Figure \ref{fig:toy_sigloss3} by the missing foreground peak in the weighted power spectrum estimate (green).  It is also evident that our result fails to recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low.  It is this behavior which we describe as signal loss.

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss3.pdf}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss4.pdf}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of foregrounds peaks at a $k$-mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare empirically estimated inverse covariance weighting where $\textbf{C}$ is derived from the data (left), and projecting out the first eigenmode only (right). In the former case, signal loss arises from the coupling of the eigenmodes of $\widehat{\textbf{C}}$ to the data. 
% JEA - redundant
%For an empirically estimated $\widehat{\textbf{C}}$, 
%its eigenvalues differ from those of the true covariance, where 
%the coupling between EoR-dominated eigenmodes and the data can lead to loss.
%Hence, 
There is negligible signal loss when all eigenmodes besides the foreground one are no longer correlated with the data.
%assigning identical weights to all eigenmodes except the first, 
%constructing a covariance proportional to the identity and the outer product of the sinusoid eigenmode, since we are only down-weighting the FG-dominated mode in this case.
}
	\label{fig:toy_sigloss3}
\end{figure*}

As discussed in Section \ref{sec:QE}, this behavior is {\it not} expected in the case that we were to use a true $\textbf{C}^{-1}$ weighting.  Rather, we would obtain the behavior shown in the toy model in Appendix \ref{sec:icw_appendix}\footnote{Note that there the true covariance matrix is also the sum of a diagonal portion describing the signal, and a single mode describing the contaminant (similar to Figure \ref{fig:toy_sigloss12}).}, with suppression of the foreground mode resulting in a nearly unbiased estimate of the power spectrum.   The key difference is that since $\widehat{\textbf{C}}$ is estimated from the data, its eigenvectors and eigenvalues are strongly coupled to the particular data realization that was used to compute it, and this coupling leads to loss.

For the case of a mode which can be safely assumed to be predominantly a foreground, its presence in the true covariance matrix will result in the desired suppression via a kind of projection (Equations \ref{eq:IdealToyModelCovariance} and \ref{eq:IdealToyModelEstimator}); whether or not it is strongly correlated with the the actual data vector is irrelevant.  However, in the case of an empirically estimated covariance matrix, the eigenmodes of $\widehat{\textbf{C}}_{\rm EoR}$ will both be incorrect and can be correlated with the data. If these incorrect eigenmodes are not correlated with the data, it will lead to non-minimum variance estimates but will not produce the suppression of the power spectrum amplitude as seen in the left plot of Figure \ref{fig:toy_sigloss3}. As shown mathematically in Appendix \ref{sec:sigloss_appendix}, however, if $\widehat{\mathbf{C}}_{\rm EoR}$ $\textit{is}$ correlated with the data vector $\mathbf{x}$, there is a kind of projection of power in the {\it non}-foreground modes from the resulting power spectrum estimate, thus producing an estimate that is biased low.  In short, {\it if the covariance is computed from the data itself, it carries the risk of over-fitting information in the data and introducing a multiplicative bias (per $k$) to estimates of the signal.}
